\texttt{Taxonomic supertree construction with}\texttt{\emph{ Incertae sedis}}\texttt{ taxa}

### 2018-04-26 doc/incertae_sedis.tex
odd label issues fixed

In order to determine how the exclude sets should be modified to account for
\emph{incertae sedis} taxa, we first attempt to express the unmodified exclude
sets in a helpful way.  We note that a node $n$ ordinarily excludes the
descendants of $\mathcal{I}(s)$ for all of its siblings, as well as the siblings
of all of its ancestors.  This leads to the following equation:
\begin{align*}
  \mathcal{E}(n) & = \mathcal{L} - \mathcal{I}(n)\\
 & =\left[\mathcal{I}(s)\big|a\in n+ancestors(n),s\in siblings(a)\right]\\
 & =\left[\mathcal{I}(s)\big|s\in siblings(n)\right]\cup\left[\mathcal{I}(s)\big|a\in ancestors(n),s\in siblings(a)\right]\\
 & =\left[\mathcal{I}(s)\big|s\in siblings(n)\right]\cup\mathcal{E}(parent(n)) \label{eq:exclude-set-formula-traditional}
\end{align*}

When modifying the exclude set for a node $n$, we seek to find all the \emph{incertae
sedis} taxa that may intrude into it. It turns out that the incertae sedis taxa that
might intrude are just the \emph{incertae sedis} siblings of $n$ and all of its ancestors.
\begin{align*}
\mathcal{E}(n) & =\left[\mathcal{I}(s)\big|a\in n+ancestors(n),s\in siblings(a),s\text{ not \emph{incertae sedis}}\right]\\
 & =\left[\mathcal{I}(s)\big|s\in siblings(n),s\text{ not \emph{incertae sedis}}\right]\cup\\
 & \,\,\,\,\,\,\,\,\,\,\,\,\left[\mathcal{I}(s)\big|a\in ancestors(n),s\in siblings(a),s\text{ not \emph{incertae sedis}}\right]\\
 & =\left[\mathcal{I}(s)\big|s\in siblings(n),s\text{ not \emph{incertae sedis}}\right]\cup\mathcal{E}(parent(n)).\label{eq:exclude-set-formula-1a}
\end{align*}
This formula allows us to compute exclude sets via a pre-order traversal on the
taxonomy tree with the exclude set for the root node defined to be $\emptyset$.

@@

Equation (\ref{eq:exclude-set-formula-1}) leads to the following
algorithm to compute the exclude set for all nodes in a tree.
\begin{enumerate} \item Set the exclude set of the root node to be
empty \item For each \emph{node} (except the root) in preorder
\begin{itemize} \item combine the \emph{exclude} set of the parent
node with the \emph{include} set of non-\emph{incertae-sedis}
siblings.
\item store this set in a hash, with key \emph{node}
\end{itemize} \end{enumerate} This algorithm is currently implemented
in \emph{otc-solve- subproblem}.
We store the sets as \emph{std::set}.

@@

We can write the exclude set of a node $n$ in terms of
the exclude set of its parent:
\begin{align}
    \mathcal{E}(n) & = \mathcal{E}(parent(n))\cup\left[\mathcal{I}(m,n)\big|m\in
           siblings(n)\right].\label{eq:exclude-set-formula-1}
\end{align}
\subsubsection{An alternative semantics that allows placing an incertae sedis group within an incertae-sedis great-uncle}

@@

\begin{align*} 
\mathcal{E}(n) & =\mathcal{E}(parent(n))\cup\left[\mathcal{I}(m,n)\big|m\in siblings(n)\right].\label{eq:exclude-set-formula-1}
\end{align*}

>>>>

In order to determine how the exclude sets should be modified to account for
\emph{incertae sedis} taxa, we first attempt to express the unmodified exclude
sets in a helpful way.  We note that a node $n$ ordinarily excludes the
descendants of $\mathcal{I}(s)$ for all of its siblings, as well as the siblings
of all of its ancestors.  This leads to the following equation:
\begin{eqnarray*}
  \mathcal{E}(n) & = \mathcal{L} - \mathcal{I}(n)\\
 & =\left[\mathcal{I}(s)\big|a\in n+ancestors(n),s\in siblings(a)\right]\\
 & =\left[\mathcal{I}(s)\big|s\in siblings(n)\right]\cup\left[\mathcal{I}(s)\big|a\in ancestors(n),s\in siblings(a)\right]\\
 & =\left[\mathcal{I}(s)\big|s\in siblings(n)\right]\cup\mathcal{E}(parent(n)) \label{eq:exsetformtraditional}
\end{eqnarray*}

When modifying the exclude set for a node $n$, we seek to find all the \emph{incertae
sedis} taxa that may intrude into it. It turns out that the incertae sedis taxa that
might intrude are just the \emph{incertae sedis} siblings of $n$ and all of its ancestors.
\begin{align*}
\mathcal{E}(n) & =
    \left[\mathcal{I}(s)\big|a\in n+ancestors(n),s\in siblings(a),s\text{ not \emph{incertae sedis}}\right]\\
& =\left[\mathcal{I}(s)\big|s\in siblings(n),s\text{ not \emph{incertae sedis}}\right]\cup\\
& \,\,\,\,\,\,\,\,\,\,\,\,\left[\mathcal{I}(s)\big|a\in ancestors(n),s\in siblings(a),s\text{ not \emph{incertae sedis}}\right]\\
& =\left[\mathcal{I}(s)\big|s\in siblings(n),s\text{ not \emph{incertae sedis}}\right]\cup\mathcal{E}(parent(n)).
\label{eq:exsetformonea}
\end{align*}
This formula allows us to compute exclude sets via a pre-order traversal on the
taxonomy tree with the exclude set for the root node defined to be $\emptyset$.

@@

Equation (\ref{eq:exsetformone}) leads to the following
algorithm to compute the exclude set for all nodes in a tree.
\begin{enumerate} \item Set the exclude set of the root node to be
empty \item For each \emph{node} (except the root) in preorder
\begin{itemize} \item combine the \emph{exclude} set of the parent
node with the \emph{include} set of non-\emph{incertae-sedis}
siblings.
\item store this set in a hash, with key \emph{node}
\end{itemize} \end{enumerate} This algorithm is currently implemented
in \emph{otc-solve- subproblem}.
We store the sets as \emph{std::set}.

@@

We can write the exclude set of a node $n$ in terms of
the exclude set of its parent:
\begin{eqnarray*}
    \mathcal{E}(n) & = \mathcal{E}(parent(n))\cup\left[\mathcal{I}(m,n)\big|m\in
           siblings(n)\right].
           \label{eq:exsetformone}
\end{eqnarray*}
\subsubsection{An alternative semantics that allows placing an incertae sedis group within an incertae-sedis great-uncle}

@@

\begin{eqnarray*} 
\mathcal{E}(n) & =\mathcal{E}(parent(n))\cup\left[\mathcal{I}(m,n)\big|m\in siblings(n)\right].\label{eq:exsetformoneb}
\end{eqnarray*}
